Gusto: a toolkit for compatible finite element dynamical cores

### 2024-07-25 source/paper.tex
minor updates

\section{Form Manipulation Language}
\label{sec: FML}
The discussion of the time discretisation so far has been general i.e.
not specific to any particular equation set. This is also true of the
implementation of the timestepping schemes. This enables flexibility
and reproducibility (**elaborate). This functionality is enabled
through the form manipulation language (FML) which is used to
associate a label to each term in the residual form defined in the
equation class. The timestepping methods are then able to reason about
which manipulations to apply to which forms by applying filters. Here
we will describe the objects required to define labels and filters
before illustrating this with a simple example(s?).

\subsection{Form Manipulation Language Objects}

The finite element discretisation of each term in an equation is
represented in Firedrake by a \texttt{ufl.Form} object. FML defines a
\texttt{Term} object that comprises the \texttt{ufl.Form} defining the
finite element discretisation of the term and a dictionary of
key-value pairs corresponding to the FML \texttt{Label} objects that
apply to this term. An FML \texttt{Label} object has a name and an
optional value. When called on a \texttt{ufl.Form} object, the
\texttt{Label} ** Labels can also be used to tag parts of the
form. For example, the \texttt{subject} label is used to tag the
\texttt{Function} in the \texttt{ufl.Form} that represents the
prognostic variable. These operations are composable, so more than one
label can be attached to a term.

The end result is a \texttt{LabelledForm} object that holds a list of
terms, pairing individual \texttt{ufl.Form} objects with an FML
\texttt{label}. Labelled forms can be added together, multiplied by
floats and have additional terms added. More complex manipulations are
enabled by the \texttt{label\_map} method that provides a way of
filtering the terms in the form based on their labels, applying
different mappings to terms that pass or fail the conditions defined
in the filter. A common operation might be to filter for terms that
have a certain label, such as \texttt{transport}, drop terms that do
not have this label and multiply those that do by the timestep. We
shall demonstrate this in the following example.

>>>>

\section{Form Manipulation Language}
\label{sec: FML}
The discussion of the time discretisation so far has been general i.e.
not specific to any particular equation set. This is also true of the
implementation of the timestepping schemes. This enables flexibility
and reproducibility (**elaborate). This functionality is enabled
through the form manipulation language (FML) which is used to
associate a label to each term in the residual form defined in the
equation class. The timestepping methods are then able to reason about
which manipulations to apply to which forms by applying filters that
check for the presence of a label and, optionally, its value. Here we
will describe the objects required to define labels and filters before
illustrating this with a simple example(s?).

The finite element discretisation of each term in an equation is
represented in Firedrake by a \texttt{ufl.Form} object. FML defines a
\texttt{Term} object that comprises the \texttt{ufl.Form} defining the
finite element discretisation of the term and a dictionary of
key-value pairs corresponding to the FML \texttt{Label} objects that
apply to this term. An FML \texttt{Label} object has a name and an
optional value, defaulting to \texttt{True} which indicates the
presence of the label. Labels can also be used to tag parts of the
form. For example, the \texttt{subject} label is used to tag the
\texttt{Function} in the \texttt{ufl.Form} that represents the
prognostic variable. These operations are composable, so more than one
label can be attached to a term. The end result of labelling the terms
in the residual is a \texttt{LabelledForm} object that holds a list of
terms, pairing individual \texttt{ufl.Form} objects with an FML
\texttt{label}. Labelled forms can be added together, multiplied by
floats and have additional terms added. More complex manipulations are
enabled by the \texttt{label\_map} method that provides a way of
filtering the terms in the form based on their labels, applying
different mappings to terms that pass or fail the conditions defined
in the filter. A common operation might be to filter for terms that
have a certain label, such as \texttt{transport}, drop terms that do
not have this label and multiply those that do by the timestep. We
shall demonstrate this in the following example.
